A method for inference of fundamental stellar parameters using high resolution spectra

### 2014-01-15 paper/paper.tex
To run GMAur

\paragraph{SPC} A very popular method behind the Kepler follow up program is Stellar Parameter Classification (SPC) \citep{blj+12}). This compares high resolution optical echelle spectra ($R \sim$ 45,000) to high resolution synthetic grids computed with Kurucz models by a cross-correlation method. The best fit parameters and uncertainties are calculated by the relative values of the cross-correlation coefficient. It is uncertain how statistically justified this uncertainty estimate is. While this is a pixel-to-pixel method, it uses only a small region of the optical spectrum ($5000 - 5300$\AA) where the spectral library has been carefully tweaked\footnote{changes to line centers and $\log(gf)$ values were on the table} to match  high resolution models of the Sun and Vega.

Most other pixel-to-pixel techniques involve comparison via a $\chi^2$ likelihood function.

\paragraph{ULySS} ULySS \citep{kpb+09} is a mature package to make comparisons for individual spectra as well as unresolved stellar populations. Any spectral library can be used as a backend, including empirical spectral libraries, such as the one from ELODIE \citep{psk+07}. population uses an iterative kappa-sigma cleaning method to discard ``bad'' lines. They test for convergence to the same answer by starting guesses located at different parameters and seeing if the same lines are masked.

Takes into account many of the things that need to be considered, including multiplication by a polynomial and rejection of bad data or model points.

Bayesian methods exist, but they use a $\chi^2$ likelihood function \citep{sdm+07}.

\citet{sb13} mention a comprehensive Bayesian method that can fit low or high resolution spectra. They too, mention that they use a $\chi^2$ likelihood. They allude to using special masks for the spectra that block noisy and uninformative regions of the spectrum, however they do not go into further detail about how these masks were identified. Also, some stellar grids focus on the most important parts of the spectrum, as identified by experts. Quoted from \citet{sb13}, grids ``sample the wavelength windows around the spectral features important for diagnostic of FGKM stars: 3850 - 4050\AA\ (Ca I lines), 4350 - 4450\AA\ (G-band, CN sensitive), 4600 - 4900\AA\ (H$\beta$), 5100 - 5300\AA\ (Mg I triplet, main gravity diagnostics), 6400 - 6640\AA (H$\alpha$), 8400 - 8800\AA\ (Ca II triplet, also used in \emph{Gaia} and in RAVE stellar survey).'' These areas can use intense scrutiny from a spectral modeller to guarantee that the models match reality.

\citet{mga13} fit BT-Settl PHOENIX models to M dwarfs. Surely these models have some serious problems with them. They use a $\chi^2$-like likelihood function, and iterate to clean bad pixels. They iterate the method against their grid of $\sim 20$ M dwarfs and track which pixels consistently have the high residuals (median residual amount $> 10$\%, or ten times their measurement error) across all stars. Presumably, this would be due to a bad opacity for a certain line or molecular feature. Then, they mask these pixels and repeat the fitting. They claim that after two iterations the masks do not change. They note that rather than selecting good and bad pixels with a binary mask, ``a more nuanced weighting scheme would be to weight each interval (pixel) according to how consistent it is with the models, or test different weighting schemes to see which gives the best agreement with the bolometric temperatures. However regions with modest agreement between the real and synthetic spectra may contain more temperature information than regions with slightly better matches, and we have no a priori information about what spectral regions are the most temperature sensitive.''

@@

The pixel-to-pixel likelihood function will evaluate the residuals in some form of 
\begin{equation}
  {\rm residual} = \frac{ {\rm data} - {\rm model}}{ {\rm noise}}
\end{equation}
where the best model is defined as one that minimizes the residuals in some sense. We will discuss other goodness of fit functions soon. Each of these quantities are functions of wavelength $\lambda$, so moving forward we will refer to them as the data spectrum $\fD(\lambda)$, the model spectrum $\fM(\lambda)$, the noise spectrum $\sigma(\lambda)$, and the residual spectrum $R(\lambda)$.

>>>>

The pixel-to-pixel likelihood function will evaluate the residuals in some form of 
\begin{equation}
  {\rm residual} = \frac{ {\rm data} - {\rm model}}{ {\rm noise}}
\end{equation}
where the best model is defined as one that minimizes the residuals in some sense. We will discuss other goodness of fit functions soon. Each of these quantities are functions of wavelength $\lambda$, so moving forward we will refer to them as the data spectrum $\fD(\lambda)$, the model spectrum $\fM(\lambda | \vt)$, the noise spectrum $\sigma(\lambda)$, and the residual spectrum $R(\lambda | \vt)$.

Most pixel-to-pixel techniques use a $\chi^2$ likelihood function and seek to minimize the square of the residuals
\begin{equation}
  \chi^2 = \sum_\lambda \left [\frac{ \fD(\lambda) - \fM(\lambda | \vstar) }{\sigma(\lambda)} \right ]^2
\end{equation}
It appears that for many techniques, $\sigma(\lambda)$ is the Poisson photon counting errors. For some data reduction pipelines, a per-pixel ``sigma spectrum'' that accounts for spectral extraction errors due to night sky line contamination or low signal to noise is available, and should be used instead. For more discussion of the noise sources beyond statistical Poisson errors, see \S\ref{sec:residuals}.

\paragraph{ULySS} ULySS \citep{kpb+09} is a mature package to make comparisons for individual spectra as well as unresolved stellar populations. Any spectral library can be used as a backend, including empirical spectral libraries such as the one from ELODIE \citep{psk+07}. They explicitly take into account some of the necessary procedures when fitting spectra, such as continuum-multiplication by a polynomial to account for flux calibration errors and rejection of bad data or model points. They use an iterative kappa-sigma clipping method to discard ``bad'' points for the fit. They test that their clipping technique converges to the same answer by starting guesses located at different parameters $\vt$ and seeing if similar lines are masked and similar best fit parameters $\vt$ are obtained.

\paragraph{\citet{mga13}} fit BT-Settl PHOENIX models to M dwarfs. Because high quality, high resolution M dwarf spectra are difficult to obtain owing to their intrinsic faintness, the development of stellar libraries for later spectral types has typically lagged behind. They use a $\chi^2$-like likelihood function and iterate to mask bad pixels. Unlike ULySS, however, they iterate the method against a sample of $\sim 20$ M dwarfs and track which pixels consistently have the high residuals (median residual amount $> 10$\%, or ten times their measurement error) across all stars. Presumably, such a consistent residual would be due to an incorrect physical constant for a certain line or molecular feature. After masking, they repeat the fitting and claim that after two iterations the masks do not change. They note that in the future, rather than selecting good and bad pixels with a binary mask, ``a more nuanced weighting scheme would be to weight each interval (pixel) according to how consistent it is with the models, or test different weighting schemes to see which gives the best agreement with the bolometric temperatures. \textbf{However regions with modest agreement between the real and synthetic spectra may contain more temperature information than regions with slightly better matches, and we have no a priori information about what spectral regions are the most temperature sensitive.}'' Thus, masking ``bad regions'' could very well destroy stellar information in the pursuit of a better ``fit.'' This is a nuanced point that masking schemes generally neglect. We aim to address later in this document.

\paragraph{``Bayesian'' methods} do exist. \citet{sdm+07} developed a hierarchical Bayesian approach that purportedly accounts for both statistical and systematic errors in the estimation of stellar parameters. I need to work through their math more carefully, but it seems as though they use a hierarchical model for the scatter between ``true'' and observed spectrum, and the unknown scatter in the relationship between $\vg$ and the $\fM$. They claim to account for systematics, but it looks like these are systematics errors due to the performance of the instrument, such as a bad flux calibration. They use infrared spectra at low resolution ($R \sim $ 1,200), and a weighted $\chi^2$ function for goodness of fit. As far as I can ascertain, the actually pixel-to-pixel fitting does not account for systematic errors in regions of the model spectrum.

\citet{sb13} present a comprehensive Bayesian method that can fit low or high resolution spectra in addition to photometry and parallax data. They too, mention that they use a $\chi^2$ likelihood. They mention using special masks for the spectra that block noisy and uninformative regions of the spectrum, however they do not go into further detail about how these masks were created or applied. Their unpublished stellar library focuses on parts of the optical spectrum which contain the most stellar information, as determined by experts. As noted by \citet{sb13}, the grids ``sample the wavelength windows around the spectral features important for diagnostic of FGKM stars: 3850 - 4050\AA\ (Ca I lines), 4350 - 4450\AA\ (G-band, CN sensitive), 4600 - 4900\AA\ (H$\beta$), 5100 - 5300\AA\ (Mg I triplet, main gravity diagnostics), 6400 - 6640\AA\ (H$\alpha$), 8400 - 8800\AA\ (Ca II triplet, also used in \emph{Gaia} and in RAVE stellar survey).''

\paragraph{SPC} A very popular method behind the Kepler follow-up program is Stellar Parameter Classification (SPC) \citep{blj+12}). This compares high resolution optical echelle spectra ($R \sim$ 45,000) to high resolution synthetic grids computed with Kurucz models by a cross-correlation method. The best fit parameters and uncertainties are calculated by the relative values of the cross-correlation coefficient. It is uncertain how statistically justified this uncertainty estimate is. While this is a pixel-to-pixel method, it uses only a small region of the optical spectrum ($5000 - 5300$\AA) where the spectral library has been carefully tweaked\footnote{changes to line centers and $\log(gf)$ values were on the table} to match  high resolution models of the Sun and Vega.

\paragraph{General trends}
Successful pixel-to-pixel methods that are successful mask spectra lines if they are doing a full spectrum fit (although this brings up the concern of \citet{mga13}). Techniques which do not mask or just use a $\chi^2$ function are limited to a small region of the spectrum where presumably their stellar models are \emph{good enough}.
By limiting themselves to These areas can use intense scrutiny from a spectral modeller to guarantee that the models match reality.

@@
